### 2017-02-12 Detector/Detector.tex
Detailed edit of detector chapter

The \acf{LHC} \cite{lhc-machine}  is a 26.7 km long synchrotron hadron accelerator and collider below the surface of the 
Franco-Swiss border near Geneva. It
is installed in the tunnel that previously hosted the Large Electron Positron accelerator \cite{lep-design}
that was operated by \acf{CERN} between 1989 and 2000.

The \ac{LHC} was designed to collide beams of protons with each other at centre-of-mass
energies of up to 14 TeV. It therefore consists of two rings in which the two beams
of protons are individually accelerated. In addition to colliding beams of protons, the \ac{LHC}
is also used for proton--lead and lead--lead collisions.
\begin{figure}[h!]
\includegraphics[width=\textwidth]{./Detector/Plots/LHC_default.jpg}
\caption[Schematic of the CERN accelerator complex.]{Schematic of the \ac{CERN} accelerator complex \cite{lhc-schematic}.}
\label{fig:lhc_schematic}
\end{figure}
The \ac{LHC} does not operate on its own: the protons are fed into
it via a chain of accelerators. Figure \ref{fig:lhc_schematic} shows a schematic
of the \ac{CERN} accelerator complex.

Protons originate from hydrogen gas, where electrons are stripped off the 
hydrogen atoms using an electric field. The protons then pass through an injector
chain which increases the energy of the protons in several steps. After the protons
are created from the hydrogen gas, they are accelerated to an energy of
50 MeV in the Linac2 linear accelerator. They are then passed on to the
\acf{PSB} which accelerates the protons until they reach an energy of 1.4 GeV.
The next accelerator in the chain, the \acf{PS}, further accelerates the protons to 25 GeV,
with the \acf{SPS} bringing the energy up to 450 GeV. When this energy has been 
reached the protons are injected into the \ac{LHC} in two counter-rotating
beams. In the \ac{LHC} ring the beams are further accelerated to the required centre-of-mass energy
by eight \ac{RF} cavities. At design operation the beams consist of 2808 proton bunches each made up 
of $\mathcal{O}(10^{11})$ protons, spaced 25ns apart. The beams are kept in circulation
using 1232 niobium-titanium superconducting dipole magnets, cooled to operate at 
a temperature of 1.9K and generating magnetic fields of up to 8.4 Tesla.
The beams collide at four points
around the \ac{LHC} ring, where the collisions are recorded by the ATLAS\cite{atlas-jinst}, 
\acs{CMS} \cite{cms-jinst}, ALICE\cite{alice-jinst}, and LHCb\cite{lhcb-jinst} detectors.

The processes the \ac{LHC} was built to study have a small cross-section
compared with the total proton-proton inelastic cross-section, as seen from figure
\ref{fig:stirling_xs}. For example, the production cross-section of the
standard model Higgs boson is 9-10 orders of magnitude smaller than the total p--p cross section.
In order to be able to study these relatively rare processes, 
the LHC operates at high instantaneous luminosity, defined as
\begin{equation}\label{eqn:CMSLHC_luminosity}
\mathcal{L} = \frac{N_b^2n_bf_{\text{rev}}\gamma}{4\pi\epsilon_n\beta^{*}}F,
\end{equation}
where $N_b$ is the number of protons per bunch, $n_b$ the number of
bunches per beam, $f_{\text{rev}}$ the revolution frequency, $\gamma$ the 
Lorentz factor, $\epsilon_n$ the normalised beam
emittance, $\beta^{*}$ the $\beta$-function at the interaction point and F a reduction
factor due to the crossing angle.

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{./Detector/Plots/crosssections2013.jpg}
\caption[Total proton-proton cross section, and the cross-sections for
several processes studied at the LHC.]{Total proton-proton cross section, and the cross-sections
for several processes studied at the LHC \cite{stirling-crosssection}.
The cross sections of these processes
are often many orders of magnitude smaller than the total p--p cross section.}
\label{fig:stirling_xs}
\end{figure}

The \ac{LHC} began its first physics run in May 2010, colliding protons at
a centre-of-mass energy of 7 TeV, and continued operation until the end of 2012. After this the 
2-year \ac{LS1} was used to upgrade the \ac{LHC} and the detectors for collisions at a centre-of-mass
energy of 13-14 TeV. In April 2015 the \ac{LHC} restarted with collisions at a centre-of-mass energy of
13 TeV. The integrated luminosity delivered to the \ac{CMS} experiment, separated by running year, is shown in
figure \ref{fig:CMSLHC_intlumi}.

\begin{figure}[h!]
\includegraphics[width=0.85\textwidth]{./Detector/Plots/int_lumi_cumulative_pp_2.pdf}
\caption[Cumulative integrated luminosity delivered to CMS
for p--p collisions at the LHC, separated by running year.]{Cumulative integrated luminosity delivered to \ac{CMS} for p--p collisions at the \ac{LHC}, separated
by running year \cite{cms-lumi-public}.}
\label{fig:CMSLHC_intlumi}
\end{figure}

The data-taking efficiency of CMS is not 100\%, for example part of the detector might be
switched off for part of the time collisions are ongoing. In 2012, the data-taking 
efficiency was 93.5\%, in 2015 it was 90.3\% and in 2016 it was 92.4 \%

From all of the data that is recorded, only a subset is used for analysis. Data is 
certified as good for use in analyses if it is known that all relevant subdetectors
were functioning correctly. In 2012 the certification efficiency was 90.4\%,
meaning all subdetectors were working well for a dataset corresponding to 19.7 fb$^{-1}$.
Due to cryogenic issues the certification efficiency with the magnet operating
at 3.8 Tesla was 60.4 \% in 2015, leading to 2.3 fb$^{-1}$ of 
analysable data for which the full detector was operational. An additional 0.6 fb$^{-1}$
was taken with the \ac{CMS} magnet switched off. For the full 2016 p--p running period
the certification efficiency was 95.4\%, resulting in 36.26 fb$^{-1}$ of data marked good
for use in analyses.

The \ac{LHC} was designed for proton-proton collisions with a nominal
peak luminosity of $10^{32}$cm$^{-2}$s$^{-1}$. In 2012 peak
luminosities of $7.7\cdot10^{33}$cm$^{-2}$s$^{-1}$ were reached, with the peak luminosity in
2015 being $5.1\cdot10^{33}$cm$^{-2}$s$^{-1}$. Since July 2016
the \ac{LHC} has been operating at instantaneous luminosities upwards of the 
design luminosity, with peak luminosities of $1.5\cdot10^{32}$cm$^{-2}$s$^{-1}$ reached
in October 2016 \cite{cms-lumi-public}.

Due to the high operating luminosity, multiple proton-proton collisions per 
bunch crossing are likely to occur. Additional interactions on top of events of 
interest are referred to as pile-up. The average number of interactions
per bunch crossing was 21 in 2012\cite{cms-lumi-public}. The average number of
interactions per bunch crossing was lower than this at the start of Run 2, with
an average of 14 interactions per bunch crossing for the data collected
during 2015 \cite{CMS-PAS-HIG-16-006}. In the data-taking period up to August 2016 the average number of interactions
per bunch crossing increased to 24 \cite{CMS-PAS-HIG-16-037}.

\section{The \acs{CMS} detector}
\label{sec:CMSLHC_CMS}
To meet the demands of the \ac{LHC} physics programme, the
 \ac{CMS} detector was designed to perform very well in searches
for physics at the TeV scale and to be able to 
function in the challenging high-luminosity environment.
The detector is 21.6 m long, 14.6 m in diameter, and weighs
12500 tonnes \cite{cms-jinst}. It consists of several subdetectors, as illustrated
in figure \ref{fig:cms_detector}.
\begin{figure}[h!]
\includegraphics[width=0.8\textwidth]{./Detector/Plots/cms.png}
\caption[A cutaway view of the CMS detector, indicating the different sub-parts of the detector.]{A cutaway view of the \ac{CMS} detector \cite{cms-jinst}, indicating the
different sub-parts of the detector.}
\label{fig:cms_detector}
\end{figure}

The silicon tracker, a cylinder \mbox{5.8 m} in length and \mbox{2.6} m in
diameter, surrounds the interaction point. The tracker is 
surrounded by the lead-tungstate \ac{ECAL}, which is enclosed
by the brass scintillator \ac{HCAL}.

The tracking and calorimeter systems are surrounded by a superconducting
solenoid, \mbox{13 m} long, \mbox{6 m} in diameter and operating at 3.8 Tesla.
Charged particles are bent in this magnetic field, allowing for precise 
measurement of their momentum. Gaseous muon detectors are embedded in the 
iron return yoke of the solenoid.

For the measurement of physical quantities, \ac{CMS} uses a coordinate
system with the origin centred at the nominal collision point
inside the experiment. The y-axis points vertically upward, the x-axis
points radially inward toward the centre of the \ac{LHC} ring. The z-axis
points along the beam direction. As such the transverse momentum and energy, 
\pT~and \ET~are measured from the x- and y- components of the momentum and energy.
The azimuthal angle $\phi$ is measured in the x-y plane with respect to the x-axis, 
the polar angle $\theta$ measured with respect to the z-axis. A coordinate
more frequently used than the polar angle is the pseudorapidity $\eta = -\ln{[\tan{\frac{\theta}{2}}]}$.
Distances in the $\eta-\phi$ plane are given as $\Delta R = \sqrt{(\Delta\eta)^2+(\Delta\phi)^2}$.

@@

\subsection{Tracker}
\label{sec:CMSLHC_CMS_tracker}
The tracker \cite{cms-jinst} is the subdetector closest to the interaction point. It
is used for accurate reconstruction of charged particle trajectories and 
the precise reconstruction of primary and secondary vertices, the latter being important for
the identification of heavy-flavour particles. This means a small impact parameter
resolution needs to be achieved, for which a highly granular system is required. 
Due to the large number of particles emerging from 
each collision, an average of 1000 per bunch crossing (every 25 ns) at LHC design operation, the 
system also needs to be able to respond quickly in order to reconstruct trajectories accurately.
At the same time, this large particle flux calls for a 
radiation-hard design that is able to survive in this harsh environment
for a long period of time. These requirements motivate the use of a silicon tracking
system.

In a silicon tracking system, the positions of particles are recorded
due to the creation of drifting electron-hole pairs.
When a charged particle passes through the silicon, an electron-hole pair
is created, which drift under an applied electric field and produce a current
that can be read out. %LEARN A BIT MORE ABOUT THIS

The tracker, providing coverage up to $|\eta|<2.5$, consists of 
several different components. A schematic of the tracking detector, indicating
the locations of these components, is given in figure \ref{fig:CMS_tracker}.
Nearest the interaction point is the pixel detector, made
of 66 million silicon pixels, each 100 $\mu$m $\times$ 150 $\mu$m in size.
The choice of this pixel size is driven by the ability to achieve the desired spatial resolution,
which is 15-20 $\mu$m in the $r-\phi$ and $z$ directions in this case. Such a
precise resolution allows for 3D vertex reconstruction. The pixel
detector consists of 3 layers of pixels in the barrel region, with 2 pixel disks 
placed in the endcap region at either side. The barrel layers sit at radii of
4.4, 7.3 and 10.2 cm and extend out to $z=\pm 26.5$ cm. The disks are placed
at $z=\pm 34.5$ cm and $z=\pm 46.5$ cm. With this layout the pixel tracker can
provide three precies position measurements along each charged particle trajectory.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./Detector/Plots/Tracker.png}
\caption[Schematic of the CMS tracker in the 
r-z plane, indicating the positions of the pixel and
strip detectors.]{Schematic of the CMS tracker in the r-z plane, indicating the
positions of the pixel and strip detectors \cite{cms-jinst}. Each line 
on the plot represents a detector module.}
\label{fig:CMS_tracker}
\end{center}
\end{figure}

Around the pixel detector the tracking system is made up of a silicon
strip detector divided into four subsystems. The first part of the
silicon strip tracker consists of the \ac{TIB} and \ac{TID}, providing four layers of
silicon strip detectors in the barrel plus three disks at both ends. These two systems
extend out towards a radius of 55 cm. The barrel layers and disks each consist
of silicon strips which are 10 cm long, $80-141~\mu$m wide and $320~\mu$m thick. The \ac{TIB} and \ac{TID}
provide four measurements of the $r-\phi$ position  with a resolution of 23-35 $\mu$m.
The \ac{TIB} and \ac{TID} are surrounded by the \ac{TOB}, which consists of six layers of silicon strip sensors extending
up to an outer radius of 116 cm and up to $z=\pm 118$ cm. The strips in the \ac{TOB} are 500~$\mu$m thick, around 25 cm long and 122-183 $\mu$m 
wide and this subdetector provides 6 measurements of $r$ and $\phi$, with a resolution
of 35-53 $\mu$m. Beyond the z-range covered by the \ac{TOB}, coverage is provided by the \ac{TEC},
consisting of nine disks of strips. In the \ac{TEC}, the strip thickness ranges from \mbox{320-500 $\mu$m} and the the strip width ranges from $97-184 \mu$m.
This part of the system provides up to nine $\phi$ measurements.

Some of the strip modules in the detector carry a second strip detector module mounted back-to-back with a stereo angle %FIXME WTF?
of 100 mrad, in order to provide measurements of the $z-$ coordinate in the barrel. This is achieved with a resolution of 230-530 $\mu$m.

The pixel tracker will be upgraded during the extended year-end technical stop in late 2016-early 2017. This upgrade is needed to
maintain the performance of the tracker as the pile-up increases over the next few years \cite{cms-pixel-upgrade}. %or just replaced?

\subsection{\acl{ECAL}}
\label{sec:CMSLHC_CMS_ecal}
The \ac{ECAL} \cite{cms-jinst} is a hermetic homogeneous calorimeter
made of nearly 76000 lead tungstate (PbWO$_4$) crystals. It has very
good energy resolution, is highly granular and has a fast response, all qualities 
which are needed to be able to detect $H\rightarrow \gamma\gamma$ decays.

The choice of lead tungstate crystals is motivated by the
radiation-hardness of the material in combination with its
short radiation length, small Moli\`ere radius and short scintillation
decay time. With a radiation length of 0.89 cm and a Moli\`ere radius of 2.2 cm,
lead tungstate crystals allow for the construction of a compact, highly 
granular \ac{ECAL}. The compactness of the \ac{ECAL} is required  
to fit inside the bore of the magnet. 
As 80\% of the light from the crystals is emitted within 25 ns, that is
within the time between two bunch crossings at design operation of the \ac{LHC},
the lead tungstate crystals satisfy the requirements imposed by the \ac{LHC} operating
parameters.

%Moliere radius: characteristic constant of a material giving scale of transverse dimenstion
%of fully contained EM showers. It is the radius of a cylinder containing on average 
%90pct of the shower's energy deposition. Related to radiation length as 
%molrad approx 0.0265X0 (Z+1.2) with Z the atomic number.
%Radiation length: mean distance over which a high energy electron
%loses all but 1/e of its energy by bremmstrahlung and 7/9 of th emean free path for pair production by 
%a high energy photon.

When a high energy electron or photon enters a crystal, it starts a
shower producing a cascade of lower energy particles. Electrons lose
energy through bremsstrahlung, with photons undergoing $\Pe^+ \Pe^-$ 
pair production. The shower continues until the photon energy drops below the 
$\Pe^+\Pe^-$ pair production threshold, and ionisation 
starts to dominate for electrons. The shower excites the atoms in the lead-tungstate
crystals, which then emit scintillation light that is converted to a current by the 
photodetectors. To record the light of the crystals, amplifying photodetectors need
to be used, with avalanche photodiodes used in the barrel region, and vacuum phototriodes employed
in the endcaps. %Because of the larger radiation levels in the endcaps
As the crystals are 25.8 radiation lengths long, most of the shower
is contained inside them.

Figure \ref{fig:CMS_ECAL} shows a schematic of the \ac{ECAL}, indicating
the subsystems it is made up of. The \ac{EB} 
covers the pseudorapidity range up to $|\eta|<1.479$, with
crystals of $0.0174 \times 0.0174$ in $\eta - \phi$. The \ac{EE}
provide coverage beyond the range of the \ac{EB}, up to $|\eta|<3.0$.
Each endcap is divided into two halves. In front of the endcaps 
the preshower detectors, covering $1.653<|\eta|<2.6$ are sampling
calorimeters whose main use is to identify neutral pions, decaying to two photons, 
in the endcaps. Additionally
the preshower detectors aid the identification of electrons and improve the resolution
of position determination. Each of the two preshower detectors consist
of two layers of lead radiators to initiate the shower, with silicon sensors
placed after each layer of radiators to measure the deposited energy. The 
total thickness of the preshower is 3 radiation lengths.

@@

To preserve the energy resolution, the crystals and
photodetectors need to be kept stable within $\pm 0.05^{\circ}$C
as the number of scintillation photons emitted by the crystals,
and the amplification of the photodetectors, are temperature dependent.
The nominal operating temperature, $18^{\circ}$C, is ensured by 
supplying the detector with water at this temperature.

The energy resolution of the \ac{ECAL} can be parameterised as
\begin{equation}\label{eqn:ecalres}
\frac{\sigma}{E} = \frac{S}{\sqrt{E}}\oplus\frac{N}{E}\oplus C,
\end{equation}
where $S$ is the stochastic term, $N$ the noise term and $C$ the constant term.
The stochastic term arises from fluctuations in lateral shower containment and 
fluctuations in the scintillation yield, the noise term due to noise from the electronics
and additional particles in the event, and the constant term comes
from calibration errors and the non-uniformity of the longitudinal response.

The values of $S$, $N$ and $C$ have been measured in an electron
test-beam, without a magnetic field or material in front of the \ac{ECAL}, as 
$S = 0.028$ GeV$^{\frac{1}{2}}$, $N = 0.12$ GeV and $C= 0.003$.
%S = 2.83 %pm 0.03%, C = 0.26% pm 0.04%

>>>>

The \acf{LHC} \cite{lhc-machine}  is a $26.7\,\kilo\metre$ long synchrotron hadron accelerator and collider below the surface of the 
Franco-Swiss border near Geneva. It
is installed in the tunnel that previously hosted the Large Electron Positron accelerator \cite{lep-design}
that was operated by \acf{CERN} between 1989 and 2000.

The \ac{LHC} was designed to collide two beams of protons %with each other 
at centre-of-mass
energies ($\sqrt{s}$) of up to $14\,\TeV$. It therefore consists of two rings in which the beams
of protons are individually accelerated. In addition to colliding beams of protons, the \ac{LHC}
is also used for proton-lead and lead-lead collisions.
\begin{figure}[h!]
\includegraphics[width=\textwidth]{./Detector/Plots/LHC_default.jpg}
\caption[Schematic of the CERN accelerator complex.]{Schematic of the \ac{CERN} accelerator complex \cite{lhc-schematic}.}
\label{fig:lhc_schematic}
\end{figure}

The \ac{LHC} does not operate on its own: the protons are fed into
it via a chain of accelerators. Figure \ref{fig:lhc_schematic} shows a schematic
of the \ac{CERN} accelerator complex.
Protons originate from hydrogen gas, where electrons are stripped off the 
hydrogen atoms using an electric field. The protons then pass through an injector
chain which increases the energy of the protons in several steps. After the protons
are created from the hydrogen gas they are accelerated to an energy of
$50\,\MeV$ in the Linac~2 linear accelerator. They are then passed on to the
\acf{PSB} which accelerates the protons until they reach an energy of $1.4\,\GeV$.
The next accelerator in the chain, the \acf{PS}, further accelerates the protons to $25\,\GeV$,
with the \acf{SPS} bringing the energy up to $450\,\GeV$. When this energy has been 
reached the protons are injected into the \ac{LHC} in two counter-rotating
beams. In the \ac{LHC} ring the beams are further accelerated to the required centre-of-mass energy
by eight \ac{RF} cavities. At design operation the beams consist of 2808 proton bunches each made up 
of $\mathcal{O}(10^{11})$ protons, spaced $25\,\ns$ apart. The beams are kept in circulation
using 1232 niobium-titanium superconducting dipole magnets, cooled to operate at 
a temperature of $1.9\,\kelvin$ and generating magnetic fields of up to $8.4\,\tesla$.
The beams collide at four points
around the \ac{LHC} ring, where the collisions are recorded by the ATLAS\cite{atlas-jinst}, 
\acs{CMS}\cite{cms-jinst}, ALICE\cite{alice-jinst}, and LHCb\cite{lhcb-jinst} detectors.

The processes the \ac{LHC} was built to study have a small cross section
compared with the total proton-proton (p-p) inelastic cross section, as seen from figure
\ref{fig:stirling_xs}. For example, the production cross section of the
\ac{SM} Higgs boson is 9--10 orders of magnitude smaller than the total p-p cross section.
In order to be able to study these relatively rare processes  
the \ac{LHC} operates at high instantaneous luminosity, defined as
\begin{equation}\label{eqn:CMSLHC_luminosity}
\mathcal{L} = \frac{N_b^2n_bf_{\text{rev}}\gamma}{4\pi\epsilon_n\beta^{*}}F,
\end{equation}
where $N_b$ is the number of protons per bunch, $n_b$ the number of
bunches per beam, $f_{\text{rev}}$ the revolution frequency, $\gamma$ the 
Lorentz factor, $\epsilon_n$ the normalised beam
emittance, $\beta^{*}$ the $\beta$-function at the interaction point and F a reduction
factor due to the crossing angle.

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{./Detector/Plots/crosssections2013.jpg}
\caption[Total p-p cross section, and the cross sections for
several processes studied at the LHC.]{Total p-p cross section, $\sigma^{\text{tot}}$, and the cross sections
for several processes studied at the LHC as a function of $\sqrt{s}$ \cite{stirling-crosssection}.
The cross sections of these processes
are often many orders of magnitude smaller than the total p-p cross section.}
\label{fig:stirling_xs}
\end{figure}

The \ac{LHC} began its first physics run in May 2010, colliding protons at
$\sqrt{s}=7\,\TeV$, and continued operation until the end of 2012. This period is known
collectively as \ac{LHC} Run 1. After this the 
two-year \ac{LS1} was used to upgrade the \ac{LHC} and the detectors for collisions at a
$\sqrt{s}$ of $13\text{--}14\,\TeV$. In April 2015 the \ac{LHC} restarted with collisions at 
$\sqrt{s}=13\,\TeV$, marking the start of Run 2. The integrated luminosity delivered to the \ac{CMS} experiment, separated by running year, is shown in
figure \ref{fig:CMSLHC_intlumi}.

\begin{figure}[h!]
\includegraphics[width=0.85\textwidth]{./Detector/Plots/int_lumi_cumulative_pp_2.pdf}
\caption[Cumulative integrated luminosity delivered to CMS
for p-p collisions at the LHC, separated by running year.]{Cumulative integrated luminosity delivered to \ac{CMS} for p-p collisions at the \ac{LHC}, separated
by running year \cite{cms-lumi-public}.}
\label{fig:CMSLHC_intlumi}
\end{figure}

The data-taking efficiency of CMS is not 100\%, as parts of the detector might be
switched off while collisions are ongoing. In 2012 the data-taking 
efficiency was 93.5\%, in 2015 it was 90.3\% and in 2016 it was 92.4\%
From all of the data that is recorded only a subset is used for analysis. Data is 
certified as good for use in analyses if it is known that all relevant sub-detectors
were functioning correctly. In 2012 the certification efficiency was 90.4\%, 
meaning all sub-detectors were working well for a dataset corresponding to $19.7\,\invfb$. %MAYBE COME BACK TO LATER?
Due to cryogenic issues the certification efficiency with the magnet operating
at $3.8\,\tesla$ was 60.4\% in 2015, leading to $2.3\,\invfb$ of 
analysable data for which the full detector was operational. An additional $0.6\,\invfb$
was taken with the \ac{CMS} magnet switched off. For the full 2016 p-p running period
the certification efficiency was 95.4\%, resulting in $36.3\,\invfb$ of data marked good %FIXME: certification efficiency
for use in analyses.

The \ac{LHC} was designed for p-p collisions with a nominal
peak luminosity of $10^{34}\,\lumiunits$. In 2012 peak
luminosities of $7.7\times 10^{33}\,\lumiunits$ were reached, with the peak luminosity in
2015 being $5.1\times 10^{33}\,\lumiunits$. Since July 2016
the \ac{LHC} has been operating at instantaneous luminosities upwards of the 
design luminosity, with peak luminosities of $1.5\times10^{34}\,\lumiunits$ reached
in October 2016 \cite{cms-lumi-public}.

Multiple p-p collisions per bunch crossing are likely to occur due to the high
operating luminosity. 
Additional interactions on top of events of 
interest are referred to as pileup. The average number of interactions %FIXME: make clear that they are unintersting?
per bunch crossing was 21 in 2012\cite{cms-lumi-public}. The average number of
interactions per bunch crossing was lower than this at the start of Run 2, with
an average of 14 interactions per bunch crossing for the data collected
during 2015 \cite{CMS-PAS-HIG-16-006}. In the data-taking period up to August 2016 the average number of interactions
per bunch crossing increased to 24 \cite{CMS-PAS-HIG-16-037}.

\section{The \acs{CMS} detector}
\label{sec:CMSLHC_CMS}
To meet the demands of the \ac{LHC} physics programme, the
 \ac{CMS} detector was designed to have a high sensitivity in searches 
for new physics at the $\TeV$ scale and to be able to 
function in the challenging high-luminosity environment.
The detector is $21.6\,\metre$ long, $14.6\,\metre$ in diameter, and weighs
$12\,500$ tonnes \cite{cms-jinst}. It consists of several sub-detectors, arranged parallel
to the beam axis in the barrel region and perpendicular to it in the endcaps, as illustrated
in figure \ref{fig:cms_detector}.
\begin{figure}[h!]
\includegraphics[width=0.8\textwidth]{./Detector/Plots/cms.png}
\caption[A cutaway view of the CMS detector, indicating the different sub-components of the detector.]{A cutaway view of the \ac{CMS} detector \cite{cms-jinst}, indicating the
different sub-components of the detector.}
\label{fig:cms_detector}
\end{figure}

The silicon tracker, a cylinder $5.8\,\metre$ in length and $2.6\,\metre$ in
diameter, surrounds the interaction point. The tracker is 
surrounded by the lead-tungstate \ac{ECAL}, which is enclosed
by the brass scintillator \ac{HCAL}.
The tracking and calorimeter systems are surrounded by a superconducting
solenoid, $13\,\metre$ long, $6\,\metre$ in diameter and operating at $3.8\,\tesla$.
Charged particles are bent in this magnetic field, allowing for precise 
measurement of their momentum. Gaseous muon detectors are embedded in the 
iron return yoke of the solenoid.

For the measurement of physical quantities \ac{CMS} uses a coordinate
system with the origin centred at the nominal collision point
inside the experiment. The $y$ axis points vertically upward, the $x$ axis
points radially inward toward the centre of the \ac{LHC} ring. The $z$ axis
points along the beam direction. As such the transverse momentum and energy, 
\pT~and \ET, are defined by the $x$- and $y$-components of the momentum and energy.
The azimuthal angle $\phi$ is measured in the $x$-$y$ plane with respect to the $x$ axis  
and the polar angle $\theta$ is measured with respect to the $z$ axis. A coordinate
more frequently used than the polar angle is the pseudorapidity $\eta = -\ln{[\tan{\frac{\theta}{2}}]}$.%Look up: differences in eta are lorentz-invariant
Distances in the $\eta$-$\phi$ plane are given as $\Delta R = \sqrt{(\Delta\eta)^2+(\Delta\phi)^2}$.

@@

\subsection{Tracker}
\label{sec:CMSLHC_CMS_tracker}
The tracker \cite{cms-jinst} is the sub-detector closest to the interaction point. It
is used for accurate reconstruction of charged particle trajectories and 
the precise reconstruction of primary (collision) and secondary vertices, the latter being important for
the identification of the in-flight decays of heavy-flavour particles. This means a small impact parameter
resolution needs to be achieved, for which a highly granular system is required. 
Due to the large number of particles emerging from 
each collision, on average 1000 per bunch crossing at design operation, the 
system also needs to be able to respond quickly in order to reconstruct trajectories accurately.
At the same time, this large particle flux calls for a 
radiation-hard design that is able to survive in this harsh environment
for a long period of time. These requirements motivate the use of a silicon tracking
system.

In a silicon tracking system the positions of particles are recorded
via the creation of drifting electron-hole pairs.
When a charged particle passes through the silicon an electron-hole pair
is created, which drift under an applied electric field and produce a current
that can be read out. %LEARN A BIT MORE ABOUT THIS

The tracker, providing coverage up to $|\eta|<2.5$, consists of 
several different components. A schematic of the tracking detector, indicating
the locations of these components, is given in figure \ref{fig:CMS_tracker}.
The pixel detector is the closest to the interaction point.
It is composed
of 66 million silicon pixels, each $100\,\micron\,\times\,150\,\micron$ in size.
The choice of this pixel size is driven by the desired spatial resolution,
which is $15$--$20\,\micron$ in the $r$-$\phi$ and $z$~directions. Such a
precise resolution allows for 3D vertex reconstruction. The pixel
detector consists of three layers of pixels in the barrel region, with two pixel disks 
placed in the endcap region at either side. The barrel layers sit at radii of
$4.4$, $7.3$ and $10.2\,\cm$ and extend out to $z=\pm\,26.5\,\cm$. The disks are placed
at $z=\pm\,34.5\,\cm$ and $z=\pm\,46.5\cm$. With this layout the pixel tracker can
provide three precise position measurements along each charged particle trajectory.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./Detector/Plots/Tracker.png}
\caption[Schematic of the CMS tracker in the 
$r$-$z$ plane, indicating the positions of the pixel and
strip detectors.]{Schematic of the CMS tracker in the $r$-$z$ plane, indicating the
positions of the pixel and strip detectors \cite{cms-jinst}. Each line 
on the figure represents a detector module.}
\label{fig:CMS_tracker}
\end{center}
\end{figure}

Around the pixel detector the tracking system is made up of a silicon
strip detector divided into four sub-systems. The first part of the
silicon strip tracker consists of the \ac{TIB} and \ac{TID}, providing four layers of
silicon strip detectors in the barrel plus three disks at both ends. These two systems
extend out up to a radius of $55\,\cm$. The barrel layers and disks each consist
of silicon strips which are $10\,\cm$ long, $80$--$141\,\micron$ wide and $320\,\micron$ thick. The \ac{TIB} and \ac{TID}
provide four measurements of the $r$-$\phi$ position with a resolution of $23$--$35\,\micron$. 
The \ac{TIB} and \ac{TID} are surrounded by the \ac{TOB}, which consists of six layers of silicon strip sensors extending
up to an outer radius of $116\,\cm$ and up to $z=\pm\,118\,\cm$. The strips in the \ac{TOB} are $500\,\micron$ thick, 
around $25\,\cm$ long and $122$--$183\,\micron$ 
wide providing 6 measurements of $r$ and $\phi$, with a resolution
of $35$--$53\,\micron$. Beyond the $z$ range covered by the \ac{TOB}, coverage is provided by the \ac{TEC},
consisting of nine disks of strips. In the \ac{TEC} the strip thickness ranges from $320$--$500\,\micron$ and 
the the strip width ranges from $97$--$184\,\micron$. %FIXME: look up why
This part of the system provides up to nine $\phi$ measurements.

Some of the strip modules in the detector carry a second strip detector module mounted back-to-back with a stereo angle %FIXME WTF?
of $100\,$mrad, in order to provide measurements of the $z$ coordinate in the barrel. This is achieved with a resolution of $230$--$530\,\micron$.

The pixel tracker will be upgraded during the extended year-end technical stop in late 2016-early 2017. This upgrade is needed to
maintain the performance of the tracker as the pileup increases over the next few years \cite{cms-pixel-upgrade}. %or just replaced?

\subsection{\acl{ECAL}}
\label{sec:CMSLHC_CMS_ecal}
The \ac{ECAL} \cite{cms-jinst} is a hermetic homogeneous calorimeter
made of nearly $76\,000$ lead tungstate (PbWO$_4$) crystals. It has very
good energy resolution, is highly granular and has a fast response, all qualities 
which are needed to be able to detect $\PHiggslight\rightarrow \Pphoton\Pphoton$ decays.

The choice of lead tungstate crystals is motivated by the
radiation-hardness of the material in combination with its
short radiation length, small Moli\`ere radius and short scintillation
decay time. With a radiation length of $0.89\,\cm$ and a Moli\`ere radius of $2.2\,\cm$,
lead tungstate crystals allow for the construction of a compact, highly 
granular \ac{ECAL}. The \ac{ECAL} is required to be compact 
for it to fit inside the bore of the magnet. 
The lead tungstate crystals satisfy the requirements imposed by the \ac{LHC} 
operating parameters as 80\% of the light from the crystals is emitted within $25\,\ns$,
which is the time between two bunch crossings at design operation.
%Moliere radius: characteristic constant of a material giving scale of transverse dimenstion
%of fully contained EM showers. It is the radius of a cylinder containing on average 
%90pct of the shower's energy deposition. Related to radiation length as 
%molrad approx 0.0265X0 (Z+1.2) with Z the atomic number.
%Radiation length: mean distance over which a high energy electron
%loses all but 1/e of its energy by bremmstrahlung and 7/9 of th emean free path for pair production by 
%a high energy photon.

When a high energy electron or photon enters a crystal it starts a
shower producing a cascade of lower energy particles. Electrons lose
energy through bremsstrahlung, with photons undergoing $\Pe^+ \Pe^-$ 
pair production. The shower continues until the photon energy drops below the 
$\Pe^+\Pe^-$ pair production threshold, and ionisation 
starts to dominate for electrons. The shower excites the atoms in the lead-tungstate
crystals, which then emit scintillation light that is converted to a current by the 
photodetectors. To record the light of the crystals amplifying photodetectors need
to be used, with avalanche photodiodes used in the barrel region, and vacuum phototriodes employed
in the endcaps. %Because of the larger radiation levels in the endcaps
As the crystals are 25.8 radiation lengths long, most of the shower
is contained inside them.

Figure \ref{fig:CMS_ECAL} shows a schematic of the \ac{ECAL}, indicating
the sub-systems it is made up of. The \ac{EB} 
covers the pseudorapidity range up to $|\eta|<1.479$, with
crystals of $0.0174 \times 0.0174$ in $\eta$-$\phi$. The \ac{EE}
provide coverage beyond the range of the \ac{EB}, up to $|\eta|<3.0$.
Each endcap is divided into two halves. The preshower detectors, 
which are sampling calorimeters placed in front of the endcaps, cover a range of $1.653<|\eta|<2.6$.
The main use of the preshower detectors is to identifiy neutral pions decaying to two photons in the 
endcaps. Additionally,
they aid in the identification of electrons and improve the resolution
of the position determination. Each of the two preshower detectors consists
of two layers of lead radiators to initiate the shower, with silicon sensors
placed after each layer of radiators to measure the deposited energy. The 
total thickness of the preshower is three radiation lengths.

@@

To preserve the energy resolution the crystals and
photodetectors need to be kept stable within $18\pm 0.05^{\circ}$C
as the number of scintillation photons emitted by the crystals,
and the amplification of the photodetectors, are temperature dependent.
This nominal operating temperature, is ensured by 
supplying the detector with water at the same temperature.

The energy resolution of the \ac{ECAL} can be parameterised as
\begin{equation}\label{eqn:ecalres}
\frac{\sigma}{E} = \frac{S}{\sqrt{E}}\oplus\frac{N}{E}\oplus C,
\end{equation}
where $S$ is the stochastic term, $N$ the noise term and $C$ the constant term.
The stochastic term arises from fluctuations in lateral shower containment and 
fluctuations in the scintillation yield, the noise term is due to noise from the electronics
and additional particles in the event, and the constant term comes
from calibration errors and the non-uniformity of the longitudinal response.

The values of $S$, $N$ and $C$ have been measured in an electron
test-beam, without a magnetic field or material in front of the \ac{ECAL}, as 
$S = 0.028\,\GeV^{\frac{1}{2}}$, $N = 0.12\,\GeV$ and $C= 0.003$.
%S = 2.83 %pm 0.03%, C = 0.26% pm 0.04%
